%% =====================================================================
%% Skontro — Software Requirements Specification
%% Template based on ISO/IEC/IEEE 29148:2018 (supersedes IEEE 830-1998)
%% Compile with: tectonic srs.tex
%% =====================================================================

\documentclass[11pt,a4paper]{article}

%% --- Encoding and language ---------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{lmodern}
\usepackage{microtype}

%% --- Layout ------------------------------------------------------------
\usepackage[a4paper,margin=25mm,headheight=14pt]{geometry}
\usepackage{fancyhdr}
\usepackage{lastpage}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Skontro --- Software Requirements Specification}
\fancyhead[R]{\small Version 1.0}
\fancyfoot[C]{\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

\usepackage{titlesec}
\usepackage{xcolor}

\definecolor{skontroBlue}{HTML}{2563EB}
\definecolor{skontroBlueDark}{HTML}{1E40AF}
\definecolor{skontroSlate}{HTML}{1F2937}
\definecolor{skontroGrey}{HTML}{F3F4F6}
\definecolor{skontroAccent}{HTML}{6366F1}
\definecolor{skontroGreen}{HTML}{059669}
\definecolor{skontroAmber}{HTML}{D97706}
\definecolor{skontroRed}{HTML}{DC2626}

\titleformat{\section}{\Large\bfseries\color{skontroSlate}}{\thesection}{0.6em}{}
\titleformat{\subsection}{\large\bfseries\color{skontroBlueDark}}{\thesubsection}{0.5em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{skontroBlueDark}}{\thesubsubsection}{0.5em}{}

%% --- Tables and lists --------------------------------------------------
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{enumitem}
\setlist[itemize]{leftmargin=*,topsep=4pt}
\setlist[enumerate]{leftmargin=*,topsep=4pt}

%% --- Hyperlinks --------------------------------------------------------
\usepackage[colorlinks=true,
            linkcolor=skontroBlue,
            urlcolor=skontroBlue,
            citecolor=skontroAccent]{hyperref}

%% --- Code listings -----------------------------------------------------
\usepackage{listings}
\lstset{
    basicstyle=\ttfamily\footnotesize,
    breaklines=true,
    columns=fullflexible,
    keepspaces=true,
    showstringspaces=false,
    frame=single,
    framesep=4pt,
    rulecolor=\color{skontroGrey}
}

%% --- Graphics ----------------------------------------------------------
\usepackage{graphicx}
\usepackage{float}

%% --- Custom requirement environment -----------------------------------
\usepackage{tcolorbox}
\tcbuselibrary{breakable,skins}

\newtcolorbox{requirement}[3]{
    enhanced,
    breakable,
    colback=skontroGrey,
    colframe=skontroBlueDark,
    boxrule=0.6pt,
    arc=2pt,
    left=8pt, right=8pt, top=6pt, bottom=6pt,
    fonttitle=\bfseries\sffamily,
    title={\textcolor{skontroSlate}{\texttt{#1}} \hfill
           \textcolor{white}{\colorbox{skontroBlue}{\strut~#2~}} \hfill
           \textit{#3}}
}

%% --- Title block -------------------------------------------------------
\title{
    \vspace*{2cm}
    \Huge\bfseries\textcolor{skontroSlate}{Skontro} \\[0.3cm]
    \Large Software Requirements Specification \\[1cm]
    \large Version 1.0
}
\author{Zeeshan Ahmed Raja}
\date{Berlin, 2026-05-27}

%% =====================================================================
\begin{document}
%% =====================================================================

\begin{titlepage}
\maketitle
\vfill
\begin{center}
    \begin{tabular}{rl}
        \textbf{Status:} & Accepted \\
        \textbf{Project:} & Skontro --- open-source German-compliant mini-ERP \\
        \textbf{Repository:} & \url{https://github.com/skontro/skontro} \\
        \textbf{Documentation:} & \url{https://docs.skontro.dev} \\
        \textbf{License:} & MIT \\
        \textbf{Standards basis:} & ISO/IEC/IEEE 29148:2018, EN 16931, \S14 UStG \\
    \end{tabular}
\end{center}
\vspace*{2cm}
\end{titlepage}

\tableofcontents
\newpage

%% =====================================================================
\section{Introduction}
%% =====================================================================

\subsection{Purpose}

This Software Requirements Specification (SRS) defines the functional and non-functional requirements for \textbf{Skontro}, an open-source self-hostable mini-ERP for German Small and Medium Enterprises (SMEs). The document serves as the contractual basis between the Vision and Scope document (the ``why'' and ``what for whom'') and the Software Design Document (the ``how'').

The document follows ISO/IEC/IEEE 29148:2018 conventions, the current standard for software and systems engineering requirements processes (which supersedes IEEE 830-1998).

\subsection{Document Conventions}

Requirements in this document follow the format:

\begin{itemize}
    \item \textbf{Functional Requirements} are prefixed \texttt{FR-NNN} and describe \textit{what} the system shall do.
    \item \textbf{Non-functional Requirements} are prefixed \texttt{NFR-NNN} and describe quality attributes (performance, security, usability, etc.).
    \item \textbf{Interface Requirements} are prefixed \texttt{IR-NNN}.
    \item Priorities follow the MoSCoW convention: \textbf{Must}, \textbf{Should}, \textbf{Could}, \textbf{Won't (this release)}.
    \item Each requirement specifies a \textbf{Source} traceable to the Vision and Scope document or an Architecture Decision Record (ADR).
    \item Each requirement specifies one or more \textbf{Acceptance criteria} that are testable.
\end{itemize}

The keywords \textbf{shall}, \textbf{should}, \textbf{may}, and \textbf{will} follow RFC 2119 conventions.

\subsection{Intended Audience and Reading Suggestions}

\begin{itemize}
    \item \textbf{Engineers} should focus on Sections 2 and 3, especially the functional requirements catalogue and the external interface requirements.
    \item \textbf{Operations engineers} should focus on Sections 3.3 (Performance), 3.4 (Reliability), and Appendix A.
    \item \textbf{Compliance auditors} should focus on Section 3.6 (Compliance Requirements) and the references to external specifications.
    \item \textbf{Project reviewers and external readers} should read Sections 1 and 2 for context, then sample the requirements catalogue in Section 3.
\end{itemize}

\subsection{Project Scope}

For the complete project scope, see the Vision and Scope document at \url{https://docs.skontro.dev/vision/vision-and-scope/}. In brief: Skontro is an open-source mini-ERP targeting the German Mittelstand, with EN 16931-compliant e-invoicing, GoBD-aligned archival, and ML-assisted bookkeeping capabilities planned for v0.3.

\subsection{References}

\begin{enumerate}
    \item Vision and Scope --- \url{https://docs.skontro.dev/vision/vision-and-scope/}
    \item ADR 0001 --- Stack Choice
    \item EN 16931 --- European standard for electronic invoicing
    \item BMF letter, e-Rechnung guidance (most recent version)
    \item Wachstumschancengesetz, BGBl. 2024 I Nr. 108
    \item \S14 UStG, \S19 UStG --- Umsatzsteuergesetz
    \item GoBD --- BMF letter dated 28.11.2019
    \item ZUGFeRD 2.1 specification --- FeRD
    \item XRechnung specification --- KoSIT
    \item ISO/IEC/IEEE 29148:2018 --- Systems and software engineering --- Life cycle processes --- Requirements engineering
    \item RFC 2119 --- Key words for use in RFCs to indicate requirement levels
\end{enumerate}

%% =====================================================================
\section{Overall Description}
%% =====================================================================

\subsection{Product Perspective}

Skontro is a standalone, self-hostable, multi-tenant mini-ERP. It is not a module of another system. It interacts with the following external systems and roles (a system context diagram appears in the arc42 architecture document, Section 3):

\begin{itemize}
    \item \textbf{Tenant Users} via web browser (Chrome, Firefox, Safari, Edge --- latest two stable versions)
    \item \textbf{Steuerberater} (tax advisor) consuming DATEV-format exports from v0.2 onwards
    \item \textbf{Invoice Recipients} consuming PDF/A-3 ZUGFeRD invoices via email or download
    \item \textbf{S3-compatible Storage} (MinIO in development, AWS S3 or Hetzner Object Storage in production)
    \item \textbf{KoSIT Validator} (continuous integration only, used to verify generated invoices)
    \item \textbf{(Future)} SMTP provider, Sentry, Stripe, Peppol access point
\end{itemize}

\subsection{Product Functions}

At v0.1.0 launch, Skontro shall provide the following high-level functions:

\begin{itemize}
    \item User authentication, registration, and tenant onboarding
    \item Multi-tenant data isolation
    \item Role-based access control (owner, admin, accountant, viewer)
    \item Customer management (company and individual)
    \item Product and service catalog management
    \item Invoice lifecycle management (draft, issued, sent, paid, cancelled)
    \item EN 16931 / ZUGFeRD 2.1 hybrid PDF/A-3 + XML invoice generation
    \item Expense tracking with receipt upload (PDF, image formats)
    \item Dashboard with key business metrics
    \item Tenant settings management (company, tax, banking, invoicing defaults)
\end{itemize}

\subsection{User Classes and Characteristics}

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{lXccc}
\toprule
\textbf{Class} & \textbf{Description} & \textbf{Tech Level} & \textbf{Frequency} & \textbf{Importance} \\
\midrule
Tenant Owner & Business owner, primary admin & Low--Medium & Daily & Critical \\
Accountant & Internal or external bookkeeper & Medium & Daily & High \\
Viewer & Read-only access (auditor, observer) & Any & Occasional & Medium \\
External Auditor & Future: read-only audit access & High & Annual & Future \\
API Integrator & Future: programmatic access & High & Continuous & Future \\
System Administrator & Self-host operator & High & Setup + as-needed & High \\
\bottomrule
\end{tabularx}
\caption{User classes for Skontro v0.1.0.}
\end{table}

\subsection{Operating Environment}

\begin{itemize}
    \item \textbf{Server OS:} Linux (Ubuntu 24.04 LTS reference); any Docker-compatible distribution
    \item \textbf{Runtime:} Docker Engine 24+, Docker Compose v2+
    \item \textbf{Languages:} PHP 8.3, Node 20 LTS, Python 3.12 (Python only required for v0.3 ML service)
    \item \textbf{Datastores:} PostgreSQL 14+ (16 recommended), Redis 7+, S3-compatible object storage
    \item \textbf{Client:} Modern evergreen browser, latest two stable versions of Chrome/Firefox/Safari/Edge; responsive from 375px viewport
    \item \textbf{Network:} HTTPS only in production; HTTP on localhost permitted in development
\end{itemize}

\subsection{Design and Implementation Constraints}

\begin{itemize}
    \item Must use Laravel 11 and Vue 3 per \textbf{ADR 0001}
    \item Must support self-hosting via \texttt{docker compose up} per \textbf{ADR 0001}
    \item Must produce EN 16931-compliant e-invoices for all issued invoices
    \item Must archive issued invoices for 10 years per GoBD (\S147 AO)
    \item Multi-tenancy via single-database \texttt{tenant\_id} discriminator (architectural decision recorded in a future ADR)
    \item API versioning via URL prefix \texttt{/api/v1}
    \item All commits follow Conventional Commits 1.0.0
    \item License: MIT
\end{itemize}

\subsection{User Documentation}

The following user-facing documentation shall be maintained:

\begin{itemize}
    \item README at the repository root with quickstart
    \item Documentation site at \url{https://docs.skontro.dev} (MkDocs Material)
    \item API reference auto-generated from code (Laravel Scribe --- OpenAPI 3.0)
    \item Operations runbook in \texttt{docs/operations/runbook.md}
    \item In-app help text on complex flows (deferred to v0.2)
\end{itemize}

\subsection{Assumptions and Dependencies}

\begin{itemize}
    \item Postgres is available and reachable
    \item Redis is available and reachable
    \item S3-compatible storage is available and reachable
    \item Working SMTP infrastructure available (v0.2+ requirement)
    \item Customer-side Steuerberater accepts DATEV CSV (v0.2+, universal in Germany)
    \item Tenant administrator can configure DNS and TLS for self-hosted deployment (v0.1 documents the procedure)
\end{itemize}

%% =====================================================================
\section{Specific Requirements}
%% =====================================================================

\subsection{External Interface Requirements}

\subsubsection{User Interfaces}

\begin{requirement}{IR-001}{Must}{UI Framework}
\textbf{Description.} The system shall provide a web-based Single Page Application built with Vue 3.

\textbf{Rationale.} Aligns with ADR 0001 stack choice; meets the productivity targets of the project.

\textbf{Source.} ADR 0001.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item The frontend bundle is served separately from the backend API
    \item Routing is client-side using Vue Router
    \item State management uses Pinia
\end{itemize}
\end{requirement}

\begin{requirement}{IR-002}{Should}{Accessibility}
\textbf{Description.} The system should conform to WCAG 2.1 Level AA for the main workflows (login, dashboard, customer list, invoice editor, expense form).

\textbf{Rationale.} German federal accessibility regulations (BFSG) progressively extend WCAG requirements; an open-source project should set a credible baseline.

\textbf{Source.} V\&S \S1.3 BO-5.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Axe-core automated audit reports zero violations on the five main workflows
    \item All interactive elements reachable via keyboard
    \item All form fields have associated labels
    \item Color contrast meets WCAG 2.1 AA on both light and dark themes
\end{itemize}
\end{requirement}

\begin{requirement}{IR-003}{Must}{Localization}
\textbf{Description.} The system shall provide a German user interface as default and an English user interface as fallback.

\textbf{Rationale.} Target market is German Mittelstand; compliance-relevant terminology must be in German.

\textbf{Source.} V\&S \S3.2 Operating Environment.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item German is the default locale on first load
    \item Users can switch to English via a settings toggle
    \item All compliance-relevant fields (Rechnung, USt-ID, Steuernummer, GoBD, etc.) are preserved as German terms regardless of UI locale
\end{itemize}
\end{requirement}

\begin{requirement}{IR-004}{Must}{Browser Support}
\textbf{Description.} The system shall support the latest two stable versions of Chrome, Firefox, Safari, and Edge.

\textbf{Source.} V\&S \S3.2 Operating Environment.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Functional tests pass on each of Chrome, Firefox, Safari, Edge (latest two versions)
    \item No reliance on proprietary browser APIs
    \item Mobile responsive design from 375px viewport width
\end{itemize}
\end{requirement}

\subsubsection{Hardware Interfaces}

Not applicable for v0.1.0. No direct hardware integration.

\subsubsection{Software Interfaces}

\begin{requirement}{IR-005}{Must}{Database Interface}
\textbf{Description.} The system shall persist all relational data in PostgreSQL 14 or later.

\textbf{Rationale.} ADR 0001 stack choice; jsonb support is required for storing parsed XRechnung XML in v0.2.

\textbf{Source.} ADR 0001.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Migrations run cleanly against Postgres 14, 15, 16
    \item No MySQL-specific or SQLite-specific SQL is used
    \item All foreign keys explicit; no implicit relationships
\end{itemize}
\end{requirement}

\begin{requirement}{IR-006}{Must}{Cache and Queue Interface}
\textbf{Description.} The system shall use Redis 7 or later for caching, session storage, and queued job processing.

\textbf{Source.} ADR 0001.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Redis URL configurable via environment variable
    \item Application degrades gracefully if Redis is temporarily unreachable (cache miss, no crash)
\end{itemize}
\end{requirement}

\begin{requirement}{IR-007}{Must}{Object Storage Interface}
\textbf{Description.} The system shall use S3-compatible object storage for receipt and document files.

\textbf{Rationale.} S3 API is the industry standard, supported by MinIO (local development), AWS S3, Hetzner Object Storage, Backblaze B2, and others.

\textbf{Source.} ADR 0001.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Storage backend configurable via environment variables (endpoint, region, bucket, access key, secret key)
    \item All file access uses signed URLs with time-limited validity
    \item Files keyed by SHA-256 hash to enable deduplication
\end{itemize}
\end{requirement}

\begin{requirement}{IR-008}{Could}{ML Service Interface}
\textbf{Description.} The system may consume a private HTTP API exposed by a separate Python FastAPI service for machine-learning inference (receipt OCR, expense categorization, cash-flow forecasting, anomaly detection).

\textbf{Rationale.} Planned for v0.3 per V\&S \S2.2.

\textbf{Source.} V\&S \S2.2.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item The backend can call the ML service if its base URL is configured
    \item If the ML service is unreachable or returns an error, the dependent feature gracefully degrades (manual entry remains available)
\end{itemize}
\end{requirement}

\subsubsection{Communications Interfaces}

\begin{requirement}{IR-010}{Must}{REST API Protocol}
\textbf{Description.} The system shall expose a versioned REST/JSON API under the URL prefix \texttt{/api/v1}.

\textbf{Source.} (To be recorded in a future ADR on API versioning.)

\textbf{Acceptance criteria:}
\begin{itemize}
    \item All endpoints return JSON with the \texttt{Content-Type: application/json; charset=utf-8} header
    \item All endpoints use standard HTTP verbs (GET, POST, PATCH, DELETE)
    \item Error responses follow a consistent shape with \texttt{message}, \texttt{code}, and an optional \texttt{errors} object
\end{itemize}
\end{requirement}

\begin{requirement}{IR-011}{Must}{HTTPS Only in Production}
\textbf{Description.} The system shall accept HTTP only on localhost in development; production deployments shall enforce HTTPS.

\textbf{Source.} Standard security practice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Production deployment runbook configures Caddy or equivalent for TLS termination
    \item Application sends HSTS header with appropriate max-age
\end{itemize}
\end{requirement}

\subsection{Functional Requirements}

\subsubsection{Authentication and User Management}

\begin{requirement}{FR-001}{Must}{User registration}
\textbf{Description.} The system shall allow a new user to register an account by providing email, password, full name, company name, country code, and (optionally) a VAT ID.

\textbf{Rationale.} Self-service onboarding is critical to the self-hostable, open-source model.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item A successful registration creates exactly one Tenant and one User in the same database transaction
    \item The newly created User is assigned the \texttt{owner} role within the new Tenant
    \item The response includes the user resource and a Sanctum access token
    \item Duplicate email registration returns HTTP 422 with a clear error message
\end{itemize}
\end{requirement}

\begin{requirement}{FR-002}{Must}{Password policy}
\textbf{Description.} The system shall enforce a minimum password strength of 12 characters including at least one digit and one symbol on registration and password change.

\textbf{Rationale.} Modern password guidance (NIST SP 800-63B); reduces successful credential-stuffing.

\textbf{Source.} NFR-009 (Security).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Registration with a password shorter than 12 characters returns HTTP 422
    \item Registration with a password missing digits or symbols returns HTTP 422
    \item The error response identifies which constraint failed
\end{itemize}
\end{requirement}

\begin{requirement}{FR-003}{Must}{Login}
\textbf{Description.} The system shall authenticate a user via email and password, returning a Sanctum personal access token on success.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Successful login returns the user resource and an access token
    \item Failed login (wrong password, unknown email) returns HTTP 401 with a generic ``invalid credentials'' message (no user-enumeration)
    \item Login is rate-limited per IP and per email (see NFR-011)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-004}{Must}{Logout}
\textbf{Description.} The system shall provide a logout endpoint that revokes the current access token.

\textbf{Source.} Standard auth practice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Logout invalidates the current token only (other tokens for the same user remain valid)
    \item Subsequent requests with the revoked token return HTTP 401
\end{itemize}
\end{requirement}

\begin{requirement}{FR-005}{Must}{Rate limiting on auth endpoints}
\textbf{Description.} The system shall apply rate limiting of 5 requests per minute per IP on the login, registration, and password-reset endpoints.

\textbf{Source.} NFR-011.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item The 6th request in 60 seconds returns HTTP 429
    \item Response includes \texttt{Retry-After} header
\end{itemize}
\end{requirement}

\begin{requirement}{FR-006}{Must}{Token TTL}
\textbf{Description.} The system shall issue access tokens with a 30-day time-to-live.

\textbf{Source.} ADR (forthcoming) on auth strategy.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Tokens expire 30 days after issue
    \item Expired tokens return HTTP 401
\end{itemize}
\end{requirement}

\begin{requirement}{FR-007}{Must}{User profile read}
\textbf{Description.} The system shall provide an authenticated endpoint that returns the current user's profile information.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Returns id, email, full\_name, locale, role within current tenant
    \item Does not return password hash or other sensitive fields
\end{itemize}
\end{requirement}

\begin{requirement}{FR-008}{Must}{Locale preference}
\textbf{Description.} The system shall persist the user's locale preference (de or en) on the user record.

\textbf{Source.} IR-003.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item PATCH to user profile accepts locale field
    \item Subsequent API responses honor the locale for localized error messages
\end{itemize}
\end{requirement}

\begin{requirement}{FR-009}{Must}{Role assignment}
\textbf{Description.} The system shall allow tenant owners and admins to assign one of four roles (owner, admin, accountant, viewer) to other users in their tenant.

\textbf{Source.} V\&S \S2.1; ADR (forthcoming) on RBAC.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Role changes require owner or admin permission
    \item Role changes are scoped to the current tenant; cannot affect users in other tenants
    \item The change is audited (timestamp, actor, target, old role, new role)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-010}{Must}{Last-owner deletion protection}
\textbf{Description.} The system shall prevent deletion or role demotion of the last user with owner role in a tenant.

\textbf{Rationale.} Prevents a tenant from being locked out of admin access.

\textbf{Source.} ADR (forthcoming) on RBAC.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Deleting the last owner returns HTTP 409 with a clear error
    \item Demoting the last owner to any non-owner role returns HTTP 409
\end{itemize}
\end{requirement}

\subsubsection{Multi-tenancy and Data Isolation}

\begin{requirement}{FR-011}{Must}{Tenant creation on registration}
\textbf{Description.} The system shall create exactly one Tenant per successful user registration and associate the new user with that Tenant as its owner.

\textbf{Source.} V\&S \S2.1; ADR (forthcoming) on multi-tenancy.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Tenant and User creation happen in a single database transaction
    \item If User creation fails, the Tenant is rolled back
    \item A unique slug is generated for the Tenant (URL-safe, derived from company name)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-012}{Must}{Data isolation between tenants}
\textbf{Description.} The system shall prevent any tenant from reading, modifying, or deleting data belonging to another tenant.

\textbf{Source.} ADR (forthcoming) on multi-tenancy.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Every tenant-owned Eloquent model uses the BelongsToTenant trait
    \item The global scope filters all queries by the authenticated user's tenant\_id
    \item Direct ID lookup of another tenant's record returns HTTP 404 (not 403)
    \item Integration test suite includes a cross-tenant isolation test for every tenant-owned model
\end{itemize}
\end{requirement}

\begin{requirement}{FR-013}{Must}{Tenant-scoped numbering sequences}
\textbf{Description.} The system shall provide atomic per-tenant sequential numbering for customers (K-YYYY-NNNNN), invoices (R-YYYY-NNNNN), and expenses (E-YYYY-NNNNN).

\textbf{Rationale.} Each tenant maintains its own numbering; concurrent writes within a tenant must not collide.

\textbf{Source.} ADR (forthcoming) on numbering strategy.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Sequence numbers are atomic under concurrent writes (row-level lock or sequence table)
    \item Numbers reset annually based on document date year
    \item No duplicate numbers within a (tenant, document\_type, year) triple
    \item Concurrency test confirms 100 parallel inserts produce 100 distinct numbers
\end{itemize}
\end{requirement}

\begin{requirement}{FR-014}{Could}{Tenant switching}
\textbf{Description.} The system may, in a future release, allow users belonging to multiple tenants to switch between them within a single session.

\textbf{Source.} V\&S \S2.2 (deferred to v0.3+).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Not implemented in v0.1.0
\end{itemize}
\end{requirement}

\begin{requirement}{FR-015}{Must}{Tenant settings model}
\textbf{Description.} The system shall provide a settings model per tenant containing company info, tax info, banking info, and invoicing defaults.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Settings are mutable only by users with owner or admin role
    \item Settings include: company name, address; USt-ID, Steuernummer, Handelsregister; IBAN, BIC, bank name; default payment terms, default tax rate, default unit, notes\_top, notes\_bottom
    \item USt-ID validation per BMF format (DE + 9 digits, checksum)
    \item IBAN validation per ISO 13616 checksum
\end{itemize}
\end{requirement}

\subsubsection{Customer Management}

\begin{requirement}{FR-016}{Must}{Create customer}
\textbf{Description.} The system shall allow authenticated users with at least accountant role to create a customer record of type \texttt{company} or \texttt{individual}.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Required fields: type, contact\_name; for company type, additionally company\_name
    \item Optional fields: address (street, postal code, city, country\_code), email, phone, VAT ID, notes
    \item On creation, a customer number K-YYYY-NNNNN is generated atomically
    \item Returns HTTP 201 with the created customer resource
\end{itemize}
\end{requirement}

\begin{requirement}{FR-017}{Must}{Read customer}
\textbf{Description.} The system shall provide an endpoint to retrieve a single customer by ID within the current tenant.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Returns HTTP 404 if customer ID does not exist or belongs to a different tenant
    \item Returns full customer resource including all fields
\end{itemize}
\end{requirement}

\begin{requirement}{FR-018}{Must}{Update customer}
\textbf{Description.} The system shall allow authenticated users with at least accountant role to update a customer's mutable fields.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Customer number is immutable after creation
    \item All other fields mutable via PATCH
    \item Updates audited (timestamp, actor, changed fields)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-019}{Must}{Soft-delete customer}
\textbf{Description.} The system shall support soft deletion of customers (mark as deleted but retain in database).

\textbf{Rationale.} Customers may have associated invoices that must remain readable. Soft delete prevents accidental data loss.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Soft-deleted customers are excluded from list and search results by default
    \item Existing invoices for soft-deleted customers remain accessible
    \item Admins can restore soft-deleted customers
\end{itemize}
\end{requirement}

\begin{requirement}{FR-020}{Must}{List and search customers}
\textbf{Description.} The system shall provide a paginated list endpoint supporting search by name, customer number, email, and VAT ID, plus sort by created\_at, name, or customer number.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Default page size 25, max page size 100
    \item Search is case-insensitive substring match
    \item Response includes total count, current page, last page
    \item Excludes soft-deleted customers by default; \texttt{?include\_deleted=1} includes them
\end{itemize}
\end{requirement}

\begin{requirement}{FR-021}{Must}{VAT ID validation}
\textbf{Description.} The system shall validate VAT IDs against the format pattern for the customer's country code.

\textbf{Source.} EN 16931 requires a valid VAT ID on every B2B invoice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item DE VAT ID format: \texttt{DE} + 9 digits, valid checksum
    \item AT, FR, NL, IT, ES, PL, BE, DK, SE, FI VAT ID format validated per country pattern
    \item Invalid format returns HTTP 422 with field-specific error
    \item v0.1 validates format only; online VIES verification deferred to v0.2
\end{itemize}
\end{requirement}

\begin{requirement}{FR-022}{Must}{EU country support}
\textbf{Description.} The system shall support customers in all EU member states for address and VAT ID fields.

\textbf{Source.} EN 16931.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Country code field uses ISO 3166-1 alpha-2 codes (DE, AT, FR, etc.)
    \item All 27 EU member states selectable in UI
\end{itemize}
\end{requirement}

\subsubsection{Product and Service Catalog}

\begin{requirement}{FR-023}{Must}{Create product or service}
\textbf{Description.} The system shall allow authenticated users with at least accountant role to create a product or service in the catalog.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Required: name, unit\_price\_cents (integer), vat\_rate (one of 19, 7, 0)
    \item Optional: description, unit (default \texttt{St\"uck}), sku
    \item Returns HTTP 201 with created resource
\end{itemize}
\end{requirement}

\begin{requirement}{FR-024}{Must}{German VAT rate assignment}
\textbf{Description.} The system shall allow assigning one of the German VAT rates (19\%, 7\%, 0\%) to each product or service.

\textbf{Source.} \S12 (1) UStG (19\%), \S12 (2) UStG (7\%), \S4 UStG (0\%).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item VAT rate stored as integer (19, 7, 0)
    \item No other VAT rates accepted in v0.1
    \item Reverse-charge (\S13b UStG) deferred to v0.3
\end{itemize}
\end{requirement}

\begin{requirement}{FR-025}{Must}{Unit selection}
\textbf{Description.} The system shall support the following units for products and services, each mapped to a UN/ECE Recommendation 20 unit code for ZUGFeRD output: St\"uck (H87), Stunde (HUR), Kilogramm (KGM), Meter (MTR), Quadratmeter (MTK), Tag (DAY), pauschal (LS).

\textbf{Source.} EN 16931 requires UN/ECE Rec. 20 unit codes.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Unit selectable from the seven supported values
    \item Each unit maps to its UN/ECE code at ZUGFeRD generation time
\end{itemize}
\end{requirement}

\begin{requirement}{FR-026}{Must}{Pricing in cents}
\textbf{Description.} The system shall store all monetary amounts as integer cents (no floating-point).

\textbf{Rationale.} Floating-point arithmetic on monetary amounts causes rounding errors that fail EN 16931 validation.

\textbf{Source.} ADR (forthcoming) on monetary handling.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Database columns for monetary amounts use BIGINT or DECIMAL, never FLOAT or DOUBLE
    \item Brick/Money or equivalent library used in application code
    \item Unit prices displayed in EUR with 2 decimal places to users; stored as cents internally
\end{itemize}
\end{requirement}

\begin{requirement}{FR-027}{Must}{Archive product}
\textbf{Description.} The system shall support archiving of products and services (mark as inactive but retain in catalog).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Archived products excluded from product picker on new invoices
    \item Archived products remain readable; existing invoices referencing them unaffected
    \item Archived products can be unarchived
\end{itemize}
\end{requirement}

\begin{requirement}{FR-028}{Should}{Optional SKU}
\textbf{Description.} The system should allow assigning an optional SKU (stock-keeping unit) string to each product.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item SKU is optional and may be unique per tenant (not enforced as unique constraint in v0.1)
    \item SKU appears on invoice line items when present
\end{itemize}
\end{requirement}

\subsubsection{Invoicing Core}

\begin{requirement}{FR-029}{Must}{Create draft invoice}
\textbf{Description.} The system shall allow authenticated users with at least accountant role to create an invoice in \texttt{draft} state.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Required: customer\_id; at least one line item
    \item Optional: invoice\_date (default today), due\_date (default today + default\_payment\_terms), service\_period\_start, service\_period\_end, notes\_top, notes\_bottom
    \item Returns HTTP 201 with the created invoice
\end{itemize}
\end{requirement}

\begin{requirement}{FR-030}{Must}{Invoice line items}
\textbf{Description.} The system shall support adding multiple line items to an invoice. Each line item references a product or service from the catalog or contains free-form description, quantity, unit, unit price, and VAT rate.

\textbf{Source.} EN 16931 invoice structure.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Each line item has: description, quantity (decimal), unit, unit\_price\_cents, vat\_rate
    \item Line items can mix products from catalog and ad-hoc descriptions
    \item Line items immutable once invoice is issued
\end{itemize}
\end{requirement}

\begin{requirement}{FR-031}{Must}{Invoice totals calculation with line-level VAT rounding}
\textbf{Description.} The system shall calculate invoice totals by computing VAT per line item, rounding each line's VAT to two decimal places, then summing --- never by computing VAT on the document subtotal.

\textbf{Rationale.} EN 16931 Business Rule BR-CO-17 mandates line-level VAT calculation. Document-level VAT rounding produces 1--2 cent differences that fail KoSIT validation on multi-rate invoices.

\textbf{Source.} EN 16931 BR-CO-17.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item For each line: line\_net = quantity * unit\_price\_cents; line\_vat = round(line\_net * vat\_rate / 100, 2)
    \item Document totals: subtotal = sum(line\_net); total\_vat = sum(line\_vat); total = subtotal + total\_vat
    \item Per-VAT-rate subtotals exposed for ZUGFeRD XML output
    \item Unit test corpus includes a multi-rate invoice case proving line-level rounding produces a different total than document-level rounding
\end{itemize}
\end{requirement}

\begin{requirement}{FR-032}{Must}{Invoice state machine}
\textbf{Description.} The system shall enforce the following invoice state transitions: \texttt{draft} $\rightarrow$ \texttt{issued} $\rightarrow$ \texttt{sent} $\rightarrow$ (\texttt{partially\_paid} $|$ \texttt{paid}); any state except \texttt{paid} $\rightarrow$ \texttt{cancelled}.

\textbf{Source.} ADR (forthcoming) on invoice state machine.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Invalid transitions return HTTP 409 with the current state in the error
    \item State transitions are audited with timestamp and actor
    \item Issued invoices are immutable for line items (only payment status can update)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-033}{Must}{Issue invoice}
\textbf{Description.} The system shall provide an endpoint to transition a draft invoice to issued state. Issuing an invoice generates the invoice number (R-YYYY-NNNNN), sets issued\_at to current timestamp, and triggers ZUGFeRD PDF generation.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Issuing requires invoice in draft state
    \item Issuing locks line items (immutable thereafter)
    \item Issuing dispatches a job to generate the ZUGFeRD-compliant PDF/A-3 file (see FR-041 through FR-048)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-034}{Must}{Cancel invoice}
\textbf{Description.} The system shall allow cancellation of any invoice in state \texttt{issued}, \texttt{sent}, or \texttt{partially\_paid}, with a mandatory cancellation reason.

\textbf{Source.} GoBD requires reasons for state changes on financial documents.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Cancellation requires a non-empty reason string
    \item Cancelled invoices appear in lists with the \texttt{cancelled} state filter
    \item Cancelled invoices cannot be uncancelled (immutable after cancel)
    \item Cannot cancel an invoice in \texttt{paid} state
\end{itemize}
\end{requirement}

\begin{requirement}{FR-035}{Must}{Mark invoice paid}
\textbf{Description.} The system shall allow recording full or partial payment against an invoice.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Recording a payment captures: amount\_cents, payment\_date, payment\_method (cash, bank\_transfer, sepa\_direct\_debit, credit\_card, other), reference
    \item If cumulative payments equal invoice total, state transitions to \texttt{paid}
    \item If cumulative payments are less than invoice total, state transitions to \texttt{partially\_paid}
\end{itemize}
\end{requirement}

\begin{requirement}{FR-036}{Must}{Payment terms}
\textbf{Description.} The system shall support configurable payment terms expressed in days (default 14 days, settable per tenant in settings).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Tenant settings define a default payment term in days
    \item Per-invoice payment terms override the tenant default
    \item Due date computed as invoice\_date + payment\_terms\_days
\end{itemize}
\end{requirement}

\begin{requirement}{FR-037}{Should}{Service period}
\textbf{Description.} The system should allow recording an optional service period (start date, end date) on each invoice.

\textbf{Rationale.} \S14 UStG requires service period when service was rendered over a period.

\textbf{Source.} \S14 UStG.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Service period optional on invoice
    \item When present, appears on the rendered PDF and in the ZUGFeRD XML
\end{itemize}
\end{requirement}

\begin{requirement}{FR-038}{Must}{Notes top and bottom}
\textbf{Description.} The system shall support free-form note text above the line items table (\texttt{notes\_top}) and below it (\texttt{notes\_bottom}) on each invoice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Markdown not supported in v0.1; plain text only
    \item Default values configurable per tenant in settings
    \item Per-invoice values override tenant defaults
\end{itemize}
\end{requirement}

\begin{requirement}{FR-039}{Must}{Invoice numbering}
\textbf{Description.} The system shall generate sequential invoice numbers per tenant per year in the format \texttt{R-YYYY-NNNNN}, atomic against concurrent issuance.

\textbf{Source.} \S14 UStG requires unique sequential invoice numbers; FR-013.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Year prefix matches the invoice's issuance year
    \item Sequence is monotonic and dense (no gaps) within a (tenant, year) pair
    \item No duplicates under concurrent issuance (tested with 100 parallel issuance calls)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-040}{Should}{Customer default payment terms}
\textbf{Description.} The system should allow setting customer-specific default payment terms that override the tenant default but can themselves be overridden per invoice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Customer record can have a payment\_terms\_days field (nullable)
    \item When creating an invoice, the precedence is: invoice-specific > customer-specific > tenant default
\end{itemize}
\end{requirement}

\subsubsection{E-Invoicing (ZUGFeRD)}

\begin{requirement}{FR-041}{Must}{Generate ZUGFeRD XML}
\textbf{Description.} The system shall generate a ZUGFeRD 2.1 EN16931-profile XML representation of each issued invoice, conformant to the EN 16931 European e-invoicing standard.

\textbf{Source.} EN 16931; Wachstumschancengesetz mandate.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item XML structure follows ZUGFeRD 2.1 EN16931 profile schema
    \item All \S14 UStG mandatory fields are present (see FR-043)
    \item All UN/ECE Rec. 20 unit codes are correct per FR-025
    \item Per-line VAT calculation per BR-CO-17 (FR-031)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-042}{Must}{Generate PDF/A-3 hybrid file}
\textbf{Description.} The system shall produce a PDF/A-3 conformant file containing both the human-readable invoice rendering and the embedded \texttt{factur-x.xml} attachment.

\textbf{Source.} ZUGFeRD 2.1 specification.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Output file passes verapdf PDF/A-3 conformance check
    \item Embedded attachment is named exactly \texttt{factur-x.xml}
    \item Attachment AFRelationship is set to \texttt{Source}
    \item XMP metadata includes the ConformanceLevel marker
\end{itemize}
\end{requirement}

\begin{requirement}{FR-043}{Must}{\S14 UStG mandatory fields}
\textbf{Description.} The system shall include all fields required by \S14 UStG on every issued invoice (both in the PDF rendering and in the embedded XML).

\textbf{Source.} \S14 UStG.

\textbf{Acceptance criteria.} Every issued invoice must include:
\begin{itemize}
    \item Full name and address of the issuing business
    \item Full name and address of the recipient
    \item Steuernummer or USt-ID of the issuer
    \item Invoice date (Rechnungsdatum)
    \item Invoice number (unique, sequential)
    \item Description and quantity of goods or services
    \item Service date or service period
    \item Net amount per VAT rate, applicable VAT rate, VAT amount, gross amount
    \item Any agreed price reductions (Skonto)
    \item If applicable: reference to reverse-charge (\S13b UStG)
    \item If applicable: reference to Kleinunternehmer status (\S19 UStG, v0.2)
\end{itemize}
\end{requirement}

\begin{requirement}{FR-044}{Must}{KoSIT validation pass}
\textbf{Description.} The system shall produce ZUGFeRD invoices that pass validation against the KoSIT validator (the official German XRechnung / ZUGFeRD reference validator).

\textbf{Source.} V\&S \S1.3 BO-2.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item CI pipeline runs the KoSIT validator against a corpus of at least 10 test invoices covering edge cases (single-line, multi-rate, foreign customer, no service period, large quantities)
    \item All test corpus invoices return zero validation errors
    \item KoSIT pass rate $\geq$ 99\% per V\&S success metric
\end{itemize}
\end{requirement}

\begin{requirement}{FR-045}{Must}{UN/ECE Rec. 20 unit code mapping}
\textbf{Description.} The system shall map internal unit identifiers to the corresponding UN/ECE Recommendation 20 unit codes in the ZUGFeRD XML output (see FR-025).

\textbf{Source.} EN 16931 requires UN/ECE codes.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Mapping table: St\"uck $\rightarrow$ H87, Stunde $\rightarrow$ HUR, Kilogramm $\rightarrow$ KGM, Meter $\rightarrow$ MTR, Quadratmeter $\rightarrow$ MTK, Tag $\rightarrow$ DAY, pauschal $\rightarrow$ LS
    \item All seven mappings tested
\end{itemize}
\end{requirement}

\begin{requirement}{FR-046}{Must}{Multi-VAT-rate handling}
\textbf{Description.} The system shall correctly aggregate line items by VAT rate and output per-rate subtotals in the ZUGFeRD XML.

\textbf{Source.} EN 16931 BR-CO-17 and BR-CO-18.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item For each distinct VAT rate appearing on an invoice, the XML contains a separate \texttt{ApplicableTradeTax} block with the correct net amount, rate, and VAT amount
    \item Test corpus includes a three-line invoice mixing 19\% and 7\% rates
\end{itemize}
\end{requirement}

\begin{requirement}{FR-047}{Must}{Generation pipeline}
\textbf{Description.} The system shall implement an invoice generation pipeline: invoice data $\rightarrow$ InvoiceCalculator $\rightarrow$ Blade template $\rightarrow$ Browsershot (Chromium) $\rightarrow$ Ghostscript (PDF/A-3 conversion with sRGB ICC profile) $\rightarrow$ horstoeko/zugferd (embed factur-x.xml) $\rightarrow$ final hybrid file.

\textbf{Source.} ADR (forthcoming) on PDF generation pipeline.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Pipeline runs asynchronously as a queued job
    \item Each step has a typed exception class for failure modes
    \item Generation time p95 below 3 seconds on reference hardware
\end{itemize}
\end{requirement}

\begin{requirement}{FR-048}{Should}{Hybrid file download}
\textbf{Description.} The system should provide an authenticated download endpoint returning the hybrid PDF/A-3 + XML file for an issued invoice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Endpoint returns 404 for draft invoices (no generated PDF yet)
    \item Endpoint returns the file with \texttt{Content-Type: application/pdf} and correct \texttt{Content-Disposition} header
    \item Filename follows pattern: \texttt{Rechnung-R-YYYY-NNNNN.pdf}
\end{itemize}
\end{requirement}

\subsubsection{Expense Tracking}

\begin{requirement}{FR-049}{Must}{Create expense}
\textbf{Description.} The system shall allow authenticated users with at least accountant role to create an expense record with vendor name, date, gross amount, VAT amount, VAT rate, category, and payment method.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Required: vendor, expense\_date, gross\_amount\_cents, vat\_rate, category
    \item Optional: notes, vat\_amount\_cents (auto-calculated if not provided)
    \item Expense number E-YYYY-NNNNN generated on creation
\end{itemize}
\end{requirement}

\begin{requirement}{FR-050}{Must}{Receipt file upload}
\textbf{Description.} The system shall support attaching one or more receipt files to an expense in formats PDF, JPG, PNG, WebP, HEIC with maximum file size 10 MB per file.

\textbf{Source.} V\&S \S2.1; GoBD requires retention of original receipts.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item MIME type validated (not just extension)
    \item Files larger than 10 MB rejected with HTTP 422
    \item Multiple receipts can be associated with a single expense
\end{itemize}
\end{requirement}

\begin{requirement}{FR-051}{Must}{SHA-256 deduplication}
\textbf{Description.} The system shall compute the SHA-256 hash of every uploaded receipt and detect duplicate uploads.

\textbf{Rationale.} Prevents the same receipt being attached twice (e.g., by accident or for fraud).

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item SHA-256 hash computed server-side on upload
    \item Storage key in S3 is the hash (no duplicate storage)
    \item If a duplicate hash is uploaded within the same tenant, the API returns a warning indicating the existing expense(s) referencing that hash
\end{itemize}
\end{requirement}

\begin{requirement}{FR-052}{Must}{Configurable expense categories}
\textbf{Description.} The system shall maintain a per-tenant list of expense categories selectable at expense creation time.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Default categories seeded for new tenants in German: B\"uromaterial, Reisekosten, Bewirtung, Software, Hardware, Telekommunikation, Werbung, Sonstiges
    \item Categories can be added, renamed, or archived (not deleted) by users with admin role
\end{itemize}
\end{requirement}

\begin{requirement}{FR-053}{Must}{Soft-delete expense}
\textbf{Description.} The system shall support soft-deletion of expenses, with the constraint that issued expenses cannot be deleted (only cancelled).

\textbf{Source.} GoBD principle of unalterability.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Soft-deleted expenses excluded from list and reporting by default
    \item Original receipt files retained in storage even after soft-delete
    \item Restore action available to users with admin role
\end{itemize}
\end{requirement}

\subsubsection{Dashboard}

\begin{requirement}{FR-054}{Must}{Revenue widgets}
\textbf{Description.} The system shall display revenue figures for the current calendar month and current year-to-date on the dashboard.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Excludes cancelled invoices
    \item Includes only issued and paid invoices
    \item Refreshes on dashboard load
\end{itemize}
\end{requirement}

\begin{requirement}{FR-055}{Must}{Open and overdue invoices widget}
\textbf{Description.} The system shall display the count and total gross value of currently open invoices and currently overdue invoices.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Open: state in \{issued, sent, partially\_paid\}, due\_date $\geq$ today
    \item Overdue: state in \{issued, sent, partially\_paid\}, due\_date $<$ today
\end{itemize}
\end{requirement}

\begin{requirement}{FR-056}{Must}{12-month timeline}
\textbf{Description.} The system shall display a 12-month rolling revenue and expense timeline as a chart.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Bar or line chart with 12 monthly buckets ending current month
    \item Revenue (green) and expenses (red) shown side-by-side or stacked
    \item Renders below 2 seconds with demo data
\end{itemize}
\end{requirement}

\begin{requirement}{FR-057}{Must}{Top customers}
\textbf{Description.} The system shall display the top 5 customers by paid revenue (current year-to-date) on the dashboard.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Sorted descending by paid revenue
    \item Shows customer name and paid amount
    \item Clicking a row opens the customer detail view
\end{itemize}
\end{requirement}

\subsubsection{Tenant Settings}

\begin{requirement}{FR-058}{Must}{Company information}
\textbf{Description.} The system shall allow users with admin role to maintain company information (name, address, country) for the tenant.

\textbf{Source.} \S14 UStG (issuer information on invoices).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item All fields editable
    \item Changes audited
    \item Changes affect new invoices only; previously-issued invoices unchanged
\end{itemize}
\end{requirement}

\begin{requirement}{FR-059}{Must}{Tax information}
\textbf{Description.} The system shall maintain tax-related identifiers per tenant: USt-ID, Steuernummer, Handelsregister entry.

\textbf{Source.} \S14 UStG requires one of USt-ID or Steuernummer on every invoice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item USt-ID validated per FR-021 format
    \item Steuernummer format-validated per state pattern
    \item Handelsregister field optional, free-text format (e.g., ``HRB 12345 Amtsgericht Berlin'')
\end{itemize}
\end{requirement}

\begin{requirement}{FR-060}{Must}{Banking information}
\textbf{Description.} The system shall maintain bank account information per tenant: IBAN, BIC, bank name.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item IBAN validated per ISO 13616 checksum
    \item BIC validated per ISO 9362 format
    \item Bank name free-text
    \item Information appears in the standard footer of issued invoices
\end{itemize}
\end{requirement}

\begin{requirement}{FR-061}{Must}{Invoicing defaults}
\textbf{Description.} The system shall maintain tenant-level invoicing defaults: default payment terms in days, default unit, default tax rate, default notes\_top, default notes\_bottom.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Defaults pre-fill the corresponding fields on new invoices
    \item Defaults editable by admin role only
\end{itemize}
\end{requirement}

\begin{requirement}{FR-062}{Should}{User management within tenant}
\textbf{Description.} The system should allow tenant admins and owners to invite new users to their tenant (v0.2), change other users' roles, and remove users from their tenant.

\textbf{Source.} V\&S \S2.1.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item v0.1 supports role change and user removal but not invitation flow (deferred to v0.2)
    \item Last-owner protection (FR-010) applies
\end{itemize}
\end{requirement}

\subsection{Performance Requirements}

\begin{requirement}{NFR-001}{Must}{API response time p50}
\textbf{Description.} The system shall respond to authenticated API requests within 200 milliseconds at the 50th percentile, measured on the reference hardware (Hetzner CX22).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Measured continuously in production via APM (e.g., Sentry Performance)
    \item Excludes ZUGFeRD generation (asynchronous)
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-002}{Should}{API response time p99}
\textbf{Description.} The system should respond to authenticated API requests within 500 milliseconds at the 99th percentile on the reference hardware.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Measured continuously
    \item Excludes long-running asynchronous operations
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-003}{Should}{Dashboard load time}
\textbf{Description.} The dashboard should fully render in under 2 seconds on demo data on the reference hardware.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Time-to-interactive measured via Lighthouse on demo deployment
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-004}{Should}{PDF generation time}
\textbf{Description.} The full ZUGFeRD invoice generation pipeline should complete in under 3 seconds at the 95th percentile for a typical multi-line invoice.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Measured in background job queue stats
    \item Test invoice has 10 line items
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-005}{Must}{Query budget}
\textbf{Description.} Index endpoints shall perform at most 5 database queries total (including N+1 prevention via eager loading).

\textbf{Source.} V\&S \S1.6 R-3.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Laravel Telescope or query log assertions in feature tests verify query count
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-006}{Should}{CI build time}
\textbf{Description.} The full CI test suite should complete within 5 minutes on standard GitHub-hosted runners.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Backend test suite + frontend test suite + linting + type-checking + docs build
    \item Parallel job execution
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-007}{Should}{Concurrent users}
\textbf{Description.} The reference deployment on Hetzner CX22 should support 100 concurrent authenticated users per tenant without exceeding the p99 response time target.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Load test with k6 or equivalent confirms target
    \item Test scenario: mixed read/write workload representative of typical bookkeeping usage
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-008}{Should}{Receipt upload throughput}
\textbf{Description.} The system should accept receipt uploads at sustained throughput of 10 concurrent uploads on the reference hardware without degrading other API performance.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Upload pipeline streams to S3-compatible storage; does not buffer entire file in PHP memory
\end{itemize}
\end{requirement}

\subsection{Software System Attributes}

\subsubsection{Security}

\begin{requirement}{NFR-009}{Must}{Password hashing}
\textbf{Description.} The system shall hash passwords using bcrypt with cost factor 12 or higher.

\textbf{Source.} OWASP password storage guidance.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item bcrypt parameter cost $\geq$ 12
    \item No plaintext passwords logged or stored
    \item Password hash never returned in API responses
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-010}{Must}{Sanctum personal access tokens}
\textbf{Description.} The system shall use Laravel Sanctum personal access tokens for API authentication, with a 30-day token TTL.

\textbf{Source.} FR-006.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Tokens stored hashed in the database
    \item Token revocation immediate on logout
    \item Expired tokens automatically pruned by scheduled job
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-011}{Must}{Rate limiting on auth}
\textbf{Description.} The system shall apply rate limiting of 5 requests per minute per IP on auth endpoints (login, registration, password reset).

\textbf{Source.} FR-005.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Implemented via Laravel rate limiting middleware
    \item 429 response with Retry-After header
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-012}{Must}{HTTPS-only in production}
\textbf{Description.} Production deployments shall enforce HTTPS and send HSTS headers with appropriate max-age.

\textbf{Source.} IR-011.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item TLS termination at reverse proxy (Caddy or equivalent)
    \item HSTS max-age 31556952 (1 year) with includeSubDomains
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-013}{Must}{No PII in logs}
\textbf{Description.} The system shall not log personally identifiable information (PII) in application logs.

\textbf{Source.} GDPR/DSGVO data minimization.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Request loggers redact authentication headers, password fields, and tokens
    \item User emails and names not included in standard request logs
    \item Sentry's data scrubbing rules configured to strip PII before transmission
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-014}{Must}{Sanitized error messages}
\textbf{Description.} Production error responses shall not expose stack traces or internal implementation details.

\textbf{Source.} OWASP error handling guidance.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item APP\_DEBUG=false in production
    \item Error responses contain only a generic message and error code
    \item Detailed errors logged to Sentry, not returned to client
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-015}{Must}{File upload validation}
\textbf{Description.} The system shall validate uploaded files via MIME type sniffing (not just extension), size limit, and reject any file matching known malicious signatures.

\textbf{Source.} FR-050; OWASP file upload guidance.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item MIME type validated server-side via file content inspection
    \item Maximum 10 MB enforced
    \item Files stored with randomized keys (not user-controlled paths) to prevent path traversal
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-016}{Must}{SQL injection prevention}
\textbf{Description.} The system shall use parameterized queries (via Eloquent or PDO) exclusively; no user input shall reach raw SQL expressions.

\textbf{Source.} OWASP injection guidance.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item PHPStan rule level 8 enforces no \texttt{DB::raw()} with user-controlled input
    \item Code review checklist includes injection check
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-017}{Must}{Content Security Policy}
\textbf{Description.} Production responses shall include a Content Security Policy header that prevents inline scripts and external resource loading except from configured origins.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item CSP header defined at reverse proxy or application middleware
    \item No \texttt{unsafe-inline} or \texttt{unsafe-eval} in script-src
    \item Reporting endpoint configured for CSP violations
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-018}{Should}{Secrets management}
\textbf{Description.} The system should load secrets (database credentials, S3 keys, SMTP credentials) exclusively from environment variables; secrets shall never appear in code or in the repository.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item \texttt{.env} file in .gitignore
    \item \texttt{.env.example} committed with placeholder values
    \item Production deploys secrets via the orchestration tool's secrets feature
\end{itemize}
\end{requirement}

\subsubsection{Usability}

\begin{requirement}{NFR-019}{Should}{WCAG 2.1 AA}
\textbf{Description.} The five main workflows shall conform to WCAG 2.1 Level AA.

\textbf{Source.} IR-002.

\textbf{Acceptance criteria.} See IR-002.
\end{requirement}

\begin{requirement}{NFR-020}{Must}{German default, English fallback}
\textbf{Description.} German shall be the default UI locale; English shall be available as a fallback locale.

\textbf{Source.} IR-003.

\textbf{Acceptance criteria.} See IR-003.
\end{requirement}

\begin{requirement}{NFR-021}{Should}{Dark mode}
\textbf{Description.} The system should provide a dark mode toggle, with the choice persisted on the user record.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Toggle accessible from the main navigation
    \item Choice persists across sessions
    \item Both modes meet color-contrast targets for WCAG 2.1 AA
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-022}{Must}{Mobile responsive}
\textbf{Description.} The user interface shall be usable on viewport widths from 375px upward.

\textbf{Source.} IR-004.

\textbf{Acceptance criteria.} Manual test on Chrome DevTools device emulation for iPhone 13 mini.
\end{requirement}

\begin{requirement}{NFR-023}{Should}{Keyboard navigation}
\textbf{Description.} All interactive elements shall be reachable and operable via keyboard alone.

\textbf{Source.} IR-002; NFR-019.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Tab order is logical
    \item Focus indicators visible on all interactive elements
    \item Keyboard shortcuts documented in the docs site
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-024}{Should}{Screen reader compatibility}
\textbf{Description.} Main workflows should be navigable using a screen reader.

\textbf{Source.} IR-002.

\textbf{Acceptance criteria.} Manual testing with NVDA on Windows.
\end{requirement}

\subsubsection{Reliability}

\begin{requirement}{NFR-025}{Must}{Daily backups}
\textbf{Description.} Production deployments shall back up the database daily, with backups retained for at least 7 days.

\textbf{Source.} V\&S \S1.5 RPO target.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Daily \texttt{pg\_dump} via cron
    \item Backups encrypted at rest
    \item Documented restore procedure in operations runbook
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-026}{Must}{Documented restore procedure}
\textbf{Description.} The operations runbook shall include a step-by-step database restore procedure verified by manual exercise.

\textbf{Source.} NFR-025.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Restore procedure tested on a staging environment with timing measured
    \item Procedure documented at \texttt{docs/operations/backup-and-restore.md}
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-027}{Should}{Production uptime}
\textbf{Description.} The production demo should maintain uptime of at least 99\% measured on a rolling 30-day window.

\textbf{Source.} V\&S \S1.5.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Monitored by Uptime Kuma or equivalent
    \item Downtime budget: $\sim$7 hours per month
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-028}{Should}{RPO and RTO}
\textbf{Description.} The system should have a recovery point objective (RPO) of 24 hours and a recovery time objective (RTO) of 4 hours.

\textbf{Source.} V\&S risks mitigation.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Backup frequency at least daily (RPO)
    \item Documented restore procedure completes in under 4 hours (RTO)
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-029}{Could}{Graceful ML-service degradation}
\textbf{Description.} When the ML service is unreachable (v0.3+), dependent features shall degrade gracefully.

\textbf{Source.} IR-008.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Manual entry remains available for any field normally pre-filled by ML
    \item A non-blocking warning informs the user that ML assistance is unavailable
\end{itemize}
\end{requirement}

\subsubsection{Maintainability}

\begin{requirement}{NFR-030}{Must}{Test coverage}
\textbf{Description.} The system shall maintain test coverage of at least 70\% line coverage and at least 60\% branch coverage.

\textbf{Source.} V\&S \S1.5.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Coverage measured by Pest (PHP) and Vitest (JS)
    \item Coverage reports published to Codecov on every push to main
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-031}{Must}{Static analysis clean}
\textbf{Description.} The backend codebase shall pass PHPStan at level 8 with no errors.

\textbf{Source.} Quality goal.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item PHPStan configured in CI
    \item Level 8 enforced
    \item No baseline file (every fix is real)
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-032}{Must}{Frontend linting clean}
\textbf{Description.} The frontend codebase shall pass ESLint with the configured ruleset and Prettier with default formatting.

\textbf{Source.} Quality goal.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item ESLint and Prettier configured in CI
    \item No warnings tolerated (CI fails on warnings)
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-033}{Must}{Conventional Commits}
\textbf{Description.} All commits shall follow Conventional Commits 1.0.0 format.

\textbf{Source.} ADR 0002 (forthcoming).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item commitlint configured via husky commit-msg hook
    \item Non-conforming commits rejected at the hook
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-034}{Must}{ADRs for architectural changes}
\textbf{Description.} All architectural decisions touching data model, authentication, deployment, dependencies, or public API shall be recorded as ADRs.

\textbf{Source.} arc42 \S9.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item ADRs stored in \texttt{docs/adr/}
    \item Pull request template enforces consideration
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-035}{Must}{Public API versioning}
\textbf{Description.} The public API shall be versioned via URL prefix; breaking changes require a major version bump with a deprecation window of at least 6 months for any deprecated endpoint.

\textbf{Source.} IR-010.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item All v1 endpoints prefixed \texttt{/api/v1}
    \item Deprecation announced in CHANGELOG and via \texttt{Deprecation} header
\end{itemize}
\end{requirement}

\subsubsection{Portability}

\begin{requirement}{NFR-036}{Must}{Docker Compose deployment}
\textbf{Description.} The system shall deploy via \texttt{docker compose up} on any host with Docker Engine 24+ and Docker Compose v2+.

\textbf{Source.} ADR 0001.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item All services defined in a single \texttt{docker-compose.yml}
    \item No Kubernetes-specific or other orchestrator-specific configuration required for the reference deployment
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-037}{Should}{Multi-architecture support}
\textbf{Description.} Docker images should be built for both amd64 and arm64 architectures.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item CI publishes multi-arch images via docker buildx
    \item Images tested on at least one arm64 host (e.g., Apple Silicon, Hetzner ARM instance)
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-038}{Must}{Postgres compatibility}
\textbf{Description.} Migrations and queries shall run cleanly on Postgres 14, 15, and 16.

\textbf{Source.} IR-005.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item CI matrix runs tests against all three Postgres versions
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-039}{Must}{No proprietary database features}
\textbf{Description.} The system shall not depend on Postgres-specific extensions beyond what is included in the standard distribution.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item No \texttt{CREATE EXTENSION} requiring non-standard packages
    \item All used extensions documented in deployment guide
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-040}{Must}{No proprietary OS or runtime dependencies}
\textbf{Description.} The system shall not require any proprietary operating system or runtime (e.g., specific cloud vendor SDKs as hard dependencies).

\textbf{Acceptance criteria:}
\begin{itemize}
    \item S3 access via the generic AWS SDK (which works against MinIO, Hetzner, AWS, etc.)
    \item No vendor lock-in code paths
\end{itemize}
\end{requirement}

\subsection{Compliance Requirements}

\begin{requirement}{NFR-041}{Must}{EN 16931 conformance}
\textbf{Description.} All issued invoices shall conform to EN 16931 European e-invoicing standard.

\textbf{Source.} Wachstumschancengesetz.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item KoSIT validator passes (FR-044)
    \item All EN 16931 mandatory business rules satisfied (BR-01 through BR-66, BR-CO-01 through BR-CO-27 as applicable)
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-042}{Must}{ZUGFeRD 2.1 profile}
\textbf{Description.} The system shall use the ZUGFeRD 2.1 EN16931 profile (not EXTENDED or MINIMUM) for generated hybrid invoices.

\textbf{Source.} ADR (forthcoming) on ZUGFeRD profile selection.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item XML schema version 2.1 declared in output
    \item Profile identifier set to \texttt{urn:cen.eu:en16931:2017}
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-043}{Must}{PDF/A-3 conformance}
\textbf{Description.} The generated hybrid invoice file shall conform to PDF/A-3 (ISO 19005-3:2012).

\textbf{Source.} ZUGFeRD 2.1 specification.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item verapdf validator passes
    \item sRGB ICC profile embedded
    \item Conformance level declared in XMP metadata
\end{itemize}
\end{requirement}

\begin{requirement}{NFR-044}{Must}{\S14 UStG mandatory fields}
\textbf{Description.} All issued invoices shall include every field required by \S14 UStG.

\textbf{Source.} FR-043.

\textbf{Acceptance criteria.} See FR-043.
\end{requirement}

\begin{requirement}{NFR-045}{Must}{10-year archival}
\textbf{Description.} Issued invoices and associated receipts shall be retained for at least 10 years per \S147 AO.

\textbf{Source.} GoBD.

\textbf{Acceptance criteria:}
\begin{itemize}
    \item Soft-deleted invoices and expenses retain their underlying files
    \item Periodic data-retention export documented in operations runbook (v0.2 implementation)
    \item Original files (receipt PDFs, invoice PDF/A-3) preserved unchanged
\end{itemize}
\end{requirement}

%% =====================================================================
\appendix
%% =====================================================================

\section{Glossary}

\begin{description}
    \item[ADR] Architecture Decision Record. A short Markdown document recording one architectural decision.
    \item[arc42] Architecture documentation template, 12 sections.
    \item[DATEV] Dominant German accounting standard and software cooperative used by Steuerberater.
    \item[Eloquent] Laravel's object-relational mapper.
    \item[EN 16931] European standard for electronic invoicing.
    \item[Factur-X] French/German hybrid PDF/A-3 + XML invoice standard, equivalent to ZUGFeRD 2.1.
    \item[GoBD] Grunds\"atze zur ordnungsm\"a\ss igen F\"uhrung und Aufbewahrung von B\"uchern, Aufzeichnungen und Unterlagen in elektronischer Form sowie zum Datenzugriff --- German digital bookkeeping rules.
    \item[Handelsregister] German commercial register; companies registered there have an HRB number.
    \item[HGB] Handelsgesetzbuch --- German Commercial Code.
    \item[Kleinunternehmer] Small business under \S19 UStG, exempted from charging VAT below a turnover threshold.
    \item[KoSIT] Koordinierungsstelle f\"ur IT-Standards; maintains the XRechnung standard and reference validator.
    \item[MoSCoW] Priority scheme: Must, Should, Could, Won't (this release).
    \item[Mittelstand] German small and medium enterprises, the economic backbone of Germany.
    \item[PDF/A-3] PDF archival format that supports embedded files (used to embed factur-x.xml).
    \item[Sanctum] Laravel's first-party authentication package for APIs and SPAs.
    \item[Skontro] German bookkeeping term: a subsidiary ledger kept alongside the main books.
    \item[Steuerberater] German tax advisor; powerful gatekeeper in SME bookkeeping decisions.
    \item[Steuernummer] German tax number, distinct from USt-ID.
    \item[USt-ID] Umsatzsteuer-Identifikationsnummer; German VAT identification number (DE + 9 digits).
    \item[UStG] Umsatzsteuergesetz --- German VAT Act.
    \item[Wachstumschancengesetz] Growth Opportunities Act 2024; introduced the e-invoicing mandate.
    \item[XRechnung] German XML invoice format, EN 16931 profile, mandated for B2G (business-to-government).
    \item[ZUGFeRD] Zentraler User Guide des Forums elektronische Rechnung Deutschland; the hybrid PDF/A-3 + XML invoice standard.
\end{description}

\section{Requirements Traceability Matrix}

This matrix maps each requirement to its source (Vision and Scope section or ADR) and tracks the implementation status. The matrix is populated as components are built; in this version (v1.0 of the SRS), most entries show \texttt{TBD} (to be determined) because implementation has not yet begun.

\begin{longtable}{lll}
\toprule
\textbf{Requirement ID} & \textbf{Source} & \textbf{Implementation Status} \\
\midrule
\endfirsthead
\toprule
\textbf{Requirement ID} & \textbf{Source} & \textbf{Implementation Status} \\
\midrule
\endhead
FR-001 & V\&S \S2.1 & TBD \\
FR-002 & NFR-009 & TBD \\
FR-003 & V\&S \S2.1 & TBD \\
FR-004 & Standard auth & TBD \\
FR-005 & NFR-011 & TBD \\
FR-006 & ADR (forthcoming) & TBD \\
FR-007 & V\&S \S2.1 & TBD \\
FR-008 & IR-003 & TBD \\
FR-009 & V\&S \S2.1 & TBD \\
FR-010 & ADR (forthcoming) & TBD \\
FR-011 & V\&S \S2.1 & TBD \\
FR-012 & ADR (forthcoming) & TBD \\
FR-013 & ADR (forthcoming) & TBD \\
FR-014 & V\&S \S2.2 & Deferred (v0.3) \\
FR-015 & V\&S \S2.1 & TBD \\
FR-016 & V\&S \S2.1 & TBD \\
FR-017 & V\&S \S2.1 & TBD \\
FR-018 & V\&S \S2.1 & TBD \\
FR-019 & V\&S \S2.1 & TBD \\
FR-020 & V\&S \S2.1 & TBD \\
FR-021 & EN 16931 & TBD \\
FR-022 & EN 16931 & TBD \\
FR-023 & V\&S \S2.1 & TBD \\
FR-024 & \S12 UStG & TBD \\
FR-025 & EN 16931 & TBD \\
FR-026 & ADR (forthcoming) & TBD \\
FR-027 & V\&S \S2.1 & TBD \\
FR-028 & V\&S \S2.1 & TBD \\
FR-029 & V\&S \S2.1 & TBD \\
FR-030 & EN 16931 & TBD \\
FR-031 & EN 16931 BR-CO-17 & TBD \\
FR-032 & ADR (forthcoming) & TBD \\
FR-033 & V\&S \S2.1 & TBD \\
FR-034 & GoBD & TBD \\
FR-035 & V\&S \S2.1 & TBD \\
FR-036 & V\&S \S2.1 & TBD \\
FR-037 & \S14 UStG & TBD \\
FR-038 & V\&S \S2.1 & TBD \\
FR-039 & \S14 UStG & TBD \\
FR-040 & V\&S \S2.1 & TBD \\
FR-041 & EN 16931 & TBD \\
FR-042 & ZUGFeRD 2.1 & TBD \\
FR-043 & \S14 UStG & TBD \\
FR-044 & V\&S \S1.3 BO-2 & TBD \\
FR-045 & EN 16931 & TBD \\
FR-046 & EN 16931 BR-CO-17 & TBD \\
FR-047 & ADR (forthcoming) & TBD \\
FR-048 & V\&S \S2.1 & TBD \\
FR-049 & V\&S \S2.1 & TBD \\
FR-050 & V\&S \S2.1, GoBD & TBD \\
FR-051 & V\&S \S2.1 & TBD \\
FR-052 & V\&S \S2.1 & TBD \\
FR-053 & GoBD & TBD \\
FR-054 & V\&S \S2.1 & TBD \\
FR-055 & V\&S \S2.1 & TBD \\
FR-056 & V\&S \S2.1 & TBD \\
FR-057 & V\&S \S2.1 & TBD \\
FR-058 & \S14 UStG & TBD \\
FR-059 & \S14 UStG & TBD \\
FR-060 & V\&S \S2.1 & TBD \\
FR-061 & V\&S \S2.1 & TBD \\
FR-062 & V\&S \S2.1 & Deferred (v0.2) \\
IR-001 & ADR 0001 & TBD \\
IR-002 & V\&S \S1.3 BO-5 & TBD \\
IR-003 & V\&S \S3.2 & TBD \\
IR-004 & V\&S \S3.2 & TBD \\
IR-005 & ADR 0001 & TBD \\
IR-006 & ADR 0001 & TBD \\
IR-007 & ADR 0001 & TBD \\
IR-008 & V\&S \S2.2 & Deferred (v0.3) \\
IR-010 & ADR (forthcoming) & TBD \\
IR-011 & Security practice & TBD \\
NFR-001 & Performance target & TBD \\
NFR-002 & Performance target & TBD \\
NFR-003 & Performance target & TBD \\
NFR-004 & Performance target & TBD \\
NFR-005 & V\&S \S1.6 R-3 & TBD \\
NFR-006 & Quality target & TBD \\
NFR-007 & V\&S \S3.2 & TBD \\
NFR-008 & V\&S \S2.1 & TBD \\
NFR-009 & OWASP & TBD \\
NFR-010 & FR-006 & TBD \\
NFR-011 & FR-005 & TBD \\
NFR-012 & IR-011 & TBD \\
NFR-013 & GDPR & TBD \\
NFR-014 & OWASP & TBD \\
NFR-015 & OWASP, FR-050 & TBD \\
NFR-016 & OWASP & TBD \\
NFR-017 & Security practice & TBD \\
NFR-018 & Security practice & TBD \\
NFR-019 & IR-002 & TBD \\
NFR-020 & IR-003 & TBD \\
NFR-021 & V\&S \S2.1 & TBD \\
NFR-022 & IR-004 & TBD \\
NFR-023 & NFR-019 & TBD \\
NFR-024 & NFR-019 & TBD \\
NFR-025 & V\&S \S1.5 & TBD \\
NFR-026 & NFR-025 & TBD \\
NFR-027 & V\&S \S1.5 & TBD \\
NFR-028 & V\&S \S1.6 & TBD \\
NFR-029 & IR-008 & Deferred (v0.3) \\
NFR-030 & V\&S \S1.5 & TBD \\
NFR-031 & Quality goal & TBD \\
NFR-032 & Quality goal & TBD \\
NFR-033 & ADR 0002 & TBD \\
NFR-034 & arc42 \S9 & TBD \\
NFR-035 & IR-010 & TBD \\
NFR-036 & ADR 0001 & TBD \\
NFR-037 & Quality goal & TBD \\
NFR-038 & IR-005 & TBD \\
NFR-039 & IR-005 & TBD \\
NFR-040 & V\&S \S3.2 & TBD \\
NFR-041 & Wachstumschancengesetz & TBD \\
NFR-042 & ADR (forthcoming) & TBD \\
NFR-043 & ZUGFeRD 2.1 & TBD \\
NFR-044 & FR-043 & TBD \\
NFR-045 & GoBD & TBD \\
\bottomrule
\end{longtable}

\section{Revision History}

\begin{table}[H]
\centering
\begin{tabular}{llll}
\toprule
\textbf{Version} & \textbf{Date} & \textbf{Author} & \textbf{Change} \\
\midrule
1.0 & 2026-05-27 & Z.~A.~Raja & Initial release. \\
\bottomrule
\end{tabular}
\end{table}

%% =====================================================================
\end{document}
